### 2021-07-07 rad-icfp-2021/unf.tex
fixing unf

There are a few notable things in this syntax. 
Every primitive is indexed by a list of types from the source language, 
which corresponds to a context being carried. 
Every constant, unary, or binary operator from Source has a corresponding $n$-ary operator !op$_{T,n}$! in UNF.
Sequential composition is denoted by !;! and !e$_{1}$;e$_{2}$! means that !e$_{1}$! should be performed, and then !e$_{2}$!.
The array operators !map2! and !reduce! have extra indices that represent well-formed terms in Source.
The language has a second kind of sequential composition, which we call passing sequential compsition $\pcomp$.
In !e$_{1}$ $\pcomp$ e$_{2}$! the idea is first that !e$_{1}$! is computed, followed by !e$_{2}$!. 
However, !e$_{1}$! outputs an argument which is not used by !e$_{2}$! and is just passed by !e$_{2}$! 
along and returned with the output of !e$_{2}$!.
The language also contains a projection !proj! which forgets some of the elements, 
and a pairing operator !pair! which takes the last pair of elements and returns the pairing of these elements.

@@

We first give a translation from our Source to Source UNF in Figure~\ref{fig:source_to_unf}, which we call $\UNFSymbol$.
As discussed above, $\UNFSymbol$ sends a constant or an operator to a primitive in Source UNF. 
This term in Source UNF carries the context of the original term in Source.
$\UNFSymbol$ sequentializes a term, mimicking a left-to-right call-by-value evaluation.
Because of this, we we're trying to return a pair, this needs to be witnessed, and this is the role of !pair!.
!e!$_{1} \pcomp$ !e!$_{2}$  is used to pass the result !e!$_{1}$ which is at the last position in the tuple through !e!$_{2}$.
It required when the coming operator represents a non-unary primitive such as $+$ or !pair!.
We want to preserve the invariant that $\UNFSymbol$(!e!) represents the term !e! which is also returning its context. 
Because !x! is not free in $\Gamma \vdash$ !let x=e$_1$in e$_2$!, we need to hide !x! after $\UNFSymbol$(!e$_2$!).
This explain the projection !proj! in $\UNFSymbol$ of a !let!. 
More generally, in !e!$_{1} \comp$ !e!$_{2}$ the return value of !e!$_{1}$ acts as a new variable for !e!$_{2}$, 
is not in the context of !e!$_{1}$, but will be returned by !e!$_{2}$.
This means we need to hide all these intermediate values to preserve the invariant,
and this explains the projection !proj! in the $\UNFSymbol$ of other terms.

@@

Next, the transformation from Target UNF to Target is presented in Figure~\ref{fig:unf_to_target}. 
We call this transformation $\UNFSymbol^{-1}$, but it is not a strict inverse of $\UNFSymbol$.
Doing $\UNFSymbol$ followed by $\UNFSymbol^{-1}$ performs some kind of ANF. 
TODO: finish

>>>>

There are a few notable things in this syntax. 
Every primitive is indexed by a list of types from the source language, 
which corresponds to a context being carried. 
Every constant, unary, or binary operator from Source has a corresponding $n$-ary operator !op$_{T,n}$! in UNF.
Sequential composition is denoted by !;! and !e$_{1}$;e$_{2}$! means that !e$_{1}$! should be performed, and then !e$_{2}$!.
The array operators !map2! and !reduce! have extra indices that represent well-formed terms in Source.

% The language has a second kind of sequential composition, which we call passing sequential compsition $\pcomp$.
% In !e$_{1}$ $\pcomp$ e$_{2}$! the idea is first that !e$_{1}$! is computed, followed by !e$_{2}$!. 
% However, !e$_{1}$! outputs an argument which is not used by !e$_{2}$! and is just passed by !e$_{2}$! 
% along and returned with the output of !e$_{2}$!.
TODO: redo

The language also contains a projection !proj! which forgets some of the elements, 
and a pairing operator !pair! which takes the last pair of elements and returns the pairing of these elements.

@@

We first give a translation from our Source to Source UNF in Figure~\ref{fig:source_to_unf}, which we call $\UNFSymbol$.
As discussed above, $\UNFSymbol$ sends a constant or an operator to a primitive in Source UNF. 
This term in Source UNF carries the context of the original term in Source.
$\UNFSymbol$ sequentializes a term, mimicking a left-to-right call-by-value evaluation.
Because of this, we we're trying to return a pair, this needs to be witnessed, and this is the role of !pair!.

% !e!$_{1} \pcomp$ !e!$_{2}$  is used to pass the result !e!$_{1}$ which is at the last position in the tuple through !e!$_{2}$.
% It required when the coming operator represents a non-unary primitive such as $+$ or !pair!.
TODO:redo

We want to preserve the invariant that $\UNFSymbol$(!e!) represents the term !e! which is also returning its context. 
Because !x! is not free in $\Gamma \vdash$ !let x=e$_1$in e$_2$!, we need to hide !x! after $\UNFSymbol$(!e$_2$!).
This explain the projection !proj! in $\UNFSymbol$ of a !let!. 
More generally, in !e!$_{1} \comp$ !e!$_{2}$ the return value of !e!$_{1}$ acts as a new variable for !e!$_{2}$, 
is not in the context of !e!$_{1}$, but will be returned by !e!$_{2}$.
This means we need to hide all these intermediate values to preserve the invariant,
and this explains the projection !proj! in the $\UNFSymbol$ of other terms.

@@

Next, the transformation from Target UNF to Target is presented in Figure~\ref{fig:unf_to_target}. 
We call this transformation $\UNFSymbol^{-1}$, but it is not a strict inverse of $\UNFSymbol$.
Doing $\UNFSymbol$ followed by $\UNFSymbol^{-1}$ performs some kind of ANF. 
TODO: finish. explain what is happening. explain carefully the context story. then alpha rename the Figure to make it correct.
